---
title: "Intro to EMA"
author: "Mathew Ling"
date: "12 July 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives

- Review the motivating questions
- Introduce you to the structure of EMA data
- Discuss the preparation of the data
- Give an overview of the Analyses

## EMA questions

- Describing state-like variables
- Relationships among state-like variables
- Relationships between state and trait-like variables

## Cross-sectional Data

![The "wide" data structure](img/WideData.svg)

## Longitudinal / Repeated Measures Data

<div align="center">
<img src="img/LongData.svg" width=960 height=720>
</div>

## EMA data

- Necessarily contains longitudinal data
- Regularly contains cross-sectional data (baseline measures)
- Both are important to research questions

-----
<div align="center">
<img src="img/EMA.svg" width=960 height=720>
</div>

## Reading, Cleaning, and Recoding 

```{r cars, echo = TRUE}
summary(cars)
```

## Combining your datasets

```{r pressure}
plot(pressure)
```

## Multi-Level Modelling (MLM)

A framework to model relationships between variables where the data are clustered (non-independent)

## Conducting a MLM
